Sweat electrolytes and CFTR activity in the era of CFTR modulators

### 2025-05-15 Christine Chan/CF Registry/cf_registry_CFRD_prediction.qmd
file paths etc.

```{r setup}
#| include: false
library(survival)
library(nlme)
library(JM)
library(dplyr)
library(tidyr)
library(arsenal)
library(knitr)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry"
)
knitr::opts_knit$set(root.dir = home_dir)
```

>>>>

```{r setup}
#| include: false
library(survival)
library(nlme)
library(JM)
library()
library(dplyr)
library(tidyr)
library(arsenal)
library(knitr)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan/CF Registry"
)
knitr::opts_knit$set(root.dir = home_dir)
```

### 2025-05-15 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
file paths etc.

```{r}
p <- prev_year_age %>%
  pivot_longer(`[10,19)`:`[31, Inf)`, names_to = "Age Group") %>%
  rename(Year = ReviewYear)
p$name <- "Prevalence of CFRD"
i <- incidence %>%
  select(
    Year, `Total Incidence Age 10-18`, `Total Incidence Age 19-30`,
    `Total Incidence Age >= 31`
  ) %>%
  pivot_longer(`Total Incidence Age 10-18`:`Total Incidence Age >= 31`,
    names_to = "Age Group"
  )
i$name <- "Incidence of CFRD"
s <- screening %>%
  filter(name == "Screened OGTT") %>%
  select(ReviewYear, age_group, value) %>%
  rename(Year = ReviewYear, `Age Group` = age_group)
s$name <- "Screening Rates for CFRD"
s$value <- s$value * 100
t <- rbind(p, i, s)
t$`Age Group` <- factor(t$`Age Group`,
  levels = c(
    "[10,19)", "[19,31)", "[31, Inf)", "Total Incidence Age >= 31",
    "Total Incidence Age 10-18", "Total Incidence Age 19-30"
  ),
  labels = c(
    "10-18", "19-30", ">= 31", ">= 31",
    "10-18", "19-30"
  )
)
t <- t[t$Year >= 2004, ]
# Figure 1
fig1a <- ggplot(
  t[t$name == "Screening Rates for CFRD", ],
  aes(x = Year, y = value, linetype = `Age Group`)
) +
  geom_line() +
  theme_bw() +
  ylab("Percentage") +
  ggtitle("Screening Rates for CFRD")
fig1b <- ggplot(
  t[t$name == "Incidence of CFRD", ],
  aes(x = Year, y = value, linetype = `Age Group`)
) +
  geom_line() +
  theme_bw() +
  ylab("Per 1,000 Person-Years") +
  ggtitle("Incidence of CFRD")
fig1c <- ggplot(
  t[t$name == "Prevalence of CFRD", ],
  aes(x = Year, y = value, linetype = `Age Group`)
) +
  geom_line() +
  theme_bw() +
  ylab("Per 1,000 Person-Years") +
  ggtitle("Prevalence of CFRD")
fig1 <- fig1a + fig1b + fig1c + plot_layout(guides = "collect")
fig1
ggsave("./Dissemination/Figures/fig1.png",
  plot = fig1, width = 12, height = 3, units = "in", dpi = 900
)
# Graphical abstract is the same but with color
fig1a <- ggplot(
  t[t$name == "Screening Rates for CFRD", ],
  aes(x = Year, y = value, color = `Age Group`)
) +
  geom_line() +
  theme_bw() +
  ylab("Percentage") +
  ggtitle("Screening Rates for CFRD")
fig1b <- ggplot(
  t[t$name == "Incidence of CFRD", ],
  aes(x = Year, y = value, color = `Age Group`)
) +
  geom_line() +
  theme_bw() +
  ylab("Per 1,000 Person-Years") +
  ggtitle("Incidence of CFRD")
fig1c <- ggplot(
  t[t$name == "Prevalence of CFRD", ],
  aes(x = Year, y = value, color = `Age Group`)
) +
  geom_line() +
  theme_bw() +
  ylab("Per 1,000 Person-Years") +
  ggtitle("Prevalence of CFRD")
fig1 <- fig1a + fig1b + fig1c + plot_layout(guides = "collect")
fig1
ggsave("./Dissemination/Figures/graphical_abstract.png",
  plot = fig1, width = 12, height = 3, units = "in", dpi = 900
)
```

>>>>

```{r}
p <- prev_year_age %>%
  pivot_longer(`[10,19)`:`[31, Inf)`, names_to = "Age Group") %>%
  rename(Year = ReviewYear)
p$name <- "Prevalence of CFRD"
i <- incidence %>%
  select(
    Year, `Total Incidence Age 10-18`, `Total Incidence Age 19-30`,
    `Total Incidence Age >= 31`
  ) %>%
  pivot_longer(`Total Incidence Age 10-18`:`Total Incidence Age >= 31`,
    names_to = "Age Group"
  )
i$name <- "Incidence of CFRD"
s <- screening %>%
  filter(name == "Screened OGTT") %>%
  select(ReviewYear, age_group, value) %>%
  rename(Year = ReviewYear, `Age Group` = age_group)
s$name <- "Screening Rates for CFRD"
s$value <- s$value * 100
t <- rbind(p, i, s)
t$`Age Group` <- factor(t$`Age Group`,
  levels = c(
    "[10,19)", "[19,31)", "[31, Inf)", "Total Incidence Age >= 31",
    "Total Incidence Age 10-18", "Total Incidence Age 19-30"
  ),
  labels = c(
    "10-18", "19-30", ">= 31", ">= 31",
    "10-18", "19-30"
  )
)
t <- t[t$Year >= 2004, ]
# Figure 1
fig1a <- ggplot(
  t[t$name == "Screening Rates for CFRD", ],
  aes(x = Year, y = value, linetype = `Age Group`)
) +
  geom_line() +
  theme_bw() +
  ylab("Percentage") +
  ggtitle("Screening Rates for CFRD")
fig1b <- ggplot(
  t[t$name == "Incidence of CFRD", ],
  aes(x = Year, y = value, linetype = `Age Group`)
) +
  geom_line() +
  theme_bw() +
  ylab("Per 1,000 Person-Years") +
  ggtitle("Incidence of CFRD")
fig1c <- ggplot(
  t[t$name == "Prevalence of CFRD", ],
  aes(x = Year, y = value, linetype = `Age Group`)
) +
  geom_line() +
  theme_bw() +
  ylab("Per 1,000 Person-Years") +
  ggtitle("Prevalence of CFRD")
fig1 <- fig1a + fig1b + fig1c + plot_layout(guides = "collect") +
  plot_annotation(title = "Figure 1: Trends in Screening, Incidence, and Prevalence of CFRD from 2003-2018, by age group ")
fig1
ggsave("./Dissemination/Figures/fig1.png",
  plot = fig1, width = 12, height = 3, units = "in", dpi = 900
)
# Graphical abstract is the same but with color
fig1a <- ggplot(
  t[t$name == "Screening Rates for CFRD", ],
  aes(x = Year, y = value, color = `Age Group`)
) +
  geom_line() +
  theme_bw() +
  ylab("Percentage") +
  ggtitle("Screening Rates for CFRD")
fig1b <- ggplot(
  t[t$name == "Incidence of CFRD", ],
  aes(x = Year, y = value, color = `Age Group`)
) +
  geom_line() +
  theme_bw() +
  ylab("Per 1,000 Person-Years") +
  ggtitle("Incidence of CFRD")
fig1c <- ggplot(
  t[t$name == "Prevalence of CFRD", ],
  aes(x = Year, y = value, color = `Age Group`)
) +
  geom_line() +
  theme_bw() +
  ylab("Per 1,000 Person-Years") +
  ggtitle("Prevalence of CFRD")
fig1 <- fig1a + fig1b + fig1c + plot_layout(guides = "collect")+
  plot_annotation(title = "US Cystic Fibrosis Foundation Patient Registry Trends in Screening, Incidence, and Prevalence of Cystic Fibrosis Related Diabetes from 2003-2018")
fig1
ggsave("./Dissemination/Figures/graphical_abstract.png",
  plot = fig1, width = 13, height = 3, units = "in", dpi = 900
)
```

### 2025-05-15 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
file paths etc.

```{r setup}
#| include: false
library(readxl)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(ggpmisc)
library(pROC)
library(rspiro)
library(childsds)
library(arsenal)
library(gt)
library(reporttools)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Documents/GitHub",
  "Linux" = "/home/timvigers/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

    "Have GCT 1 Hour Plasma or OGTT 2 Hour",

@@

```{r results='asis'}
#| label: table1a
#| tbl-cap: Participant characteristics by OGTT status
#| message: false
# Basic table formula
f <- ~ HbA1c + kwt(Age, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Weight (kg)`, "N", "Nmiss2", "medianq1q3", "range") + `Height (cm)` +
  kwt(BMI, "N", "Nmiss2", "medianq1q3", "range") + Gender + fe(Race) +
  fe(Ethnicity) + fe(`Pancreatic insufficient`) +
  fe(`Type 2 diabetes in the family - Father`) +
  fe(`Type 2 diabetes in the family - Mother`) +
  fe(`Brothers (number with diabetes)`) + fe(`Sisters (number with diabetes)`) +
  kwt(
    `Number of other relatives with diabetes (aunts, uncles, cousins, grandparents)`,
    "N", "Nmiss2", "medianq1q3", "range"
  ) +
  `Highest FEV1pp` + `Second Highest FEV1pp` + `Average of Highest FEV1pp` +
  kwt(`GCT 0 min plasma`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`GCT 0 min capillary`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`GCT 60 min plasma`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`GCT 60 min capillary`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`OGTT 0 min`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`OGTT 60 min`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`OGTT 120 min`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Height %ile`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Weight %ile`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`BMI %ile`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`BMI Value (Adults Only)`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Days From OGTT to GCT`, "N", "Nmiss2", "medianq1q3", "range") +
  fe(Diagnosis) + fe(Dysglycemia) + fe(Diabetes) + fe(CFRD)
t1a <- tableby(update(f, `Got OGTT and GCT?` ~ . - `Got OGTT and GCT?`),
  data = df, control = mycontrols
)
summary(t1a, pfootnote = T, digits = 2)
# Limit to 189 after this
df <- df %>%
  filter(`Got OGTT and GCT?` == "Have GCT 1 Hour Plasma or OGTT 2 Hour") %>%
  select(-`Got OGTT and GCT?`)
```

@@

>>>>

```{r setup}
#| include: false
library(readxl)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(ggpmisc)
library(pROC)
library(rspiro)
library(childsds)
library(arsenal)
library(gt)
library(reporttools)
library(rpart)
library(rpart.plot)
library(rms)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Documents/GitHub",
  "Linux" = "/home/timvigers/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

    "Have GCT 1 Hour Plasma and OGTT 2 Hour",

@@

```{r results='asis'}
#| label: table1a
#| tbl-cap: Participant characteristics by OGTT status
#| message: false
# Basic table formula
f <- ~ HbA1c + kwt(Age, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Weight (kg)`, "N", "Nmiss2", "medianq1q3", "range") + `Height (cm)` +
  kwt(BMI, "N", "Nmiss2", "medianq1q3", "range") + Gender + fe(Race) +
  fe(Ethnicity) + fe(`Pancreatic insufficient`) +
  fe(`Type 2 diabetes in the family - Father`) +
  fe(`Type 2 diabetes in the family - Mother`) +
  fe(`Brothers (number with diabetes)`) + fe(`Sisters (number with diabetes)`) +
  kwt(
    `Number of other relatives with diabetes (aunts, uncles, cousins, grandparents)`,
    "N", "Nmiss2", "medianq1q3", "range"
  ) +
  `Highest FEV1pp` + `Second Highest FEV1pp` + `Average of Highest FEV1pp` +
  kwt(`GCT 0 min plasma`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`GCT 0 min capillary`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`GCT 60 min plasma`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`GCT 60 min capillary`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`OGTT 0 min`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`OGTT 60 min`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`OGTT 120 min`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Height %ile`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Weight %ile`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`BMI %ile`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`BMI Value (Adults Only)`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Days From OGTT to GCT`, "N", "Nmiss2", "medianq1q3", "range") +
  fe(Diagnosis) + fe(Dysglycemia) + fe(Diabetes) + fe(CFRD)
t1a <- tableby(update(f, `Got OGTT and GCT?` ~ . - `Got OGTT and GCT?`),
  data = df, control = mycontrols
)
summary(t1a, pfootnote = T, digits = 2)
# Limit to 189 after this
df <- df %>%
  filter(`Got OGTT and GCT?` == "Have GCT 1 Hour Plasma and OGTT 2 Hour") %>%
  select(-`Got OGTT and GCT?`)
```

@@

# Nomogram

### `GCT 0 min capillary` and HbA1c

```{r}
f <- lrm(CFRD ~ `GCT 0 min capillary` + HbA1c,
  data = df
)
ddist <- datadist(df)
options(datadist = "ddist")
nom <- nomogram(f, fun = plogis, funlabel = "Risk of CFRD",lp = F)
plot(nom)
```

Here’s how to use a nomogram. Pick a point on the GCT 0 min capillary and HbA1c scales and get the associated points for each. Add the points together to get the associated probability based on the Total Points line.

For example:

- GCT 0 min capillary = 130 (40 points)
- HbA1c = 6.6 (10 points)

Add up the points:

40 + 10 = 50 Total Points

50 Total Points lines up at about 0.08, so this person has approximately an 8% chance of CFRD. The process is the same for the nomograms below.

### `GCT 0 min capillary`, HbA1c, and `GCT 60 min plasma`

```{r}
f <- lrm(CFRD ~ `GCT 0 min capillary` + HbA1c + `GCT 60 min plasma`,
  data = df
)
ddist <- datadist(df)
options(datadist = "ddist")
nom <- nomogram(f, fun = plogis, funlabel = "Risk of CFRD",lp = F)
plot(nom)
```

### `GCT 0 min capillary`, and `GCT 60 min plasma`

```{r}
f <- lrm(CFRD ~ `GCT 0 min capillary` + `GCT 60 min plasma`,
  data = df
)
ddist <- datadist(df)
options(datadist = "ddist")
nom <- nomogram(f, fun = plogis, funlabel = "Risk of CFRD",lp = F)
plot(nom)
```

# Decision trees

## Predicting CFRD

```{r}
tree <- rpart(CFRD ~ `GCT 0 min capillary` + HbA1c + `GCT 60 min plasma`,
  data = df
)
rpart.plot(tree, type = 0)
```

We just don't have enough data to do this properly for CFRD.

## Predicting dysglycemia

```{r}
tree <- rpart(Dysglycemia ~ `GCT 0 min capillary` + `GCT 60 min plasma`, data = df)
rpart.plot(tree, type = 0)
```

### 2025-05-15 Edith Zemanick/Sweat Variability/sweat_test_analysis.qmd
file paths etc.

```{r setup}
#| include: false
library(Hmisc)
library(lubridate)
library(readxl)
library(tidyverse)
library(arsenal)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Edith Zemanick/Sweat Variability"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| message: false
# Import CHCO, filter out those with no demo data, rename columns, etc.
chco <- read.csv("./Data_Cleaned/chco_reidentified.csv",
  na.strings = c(
    "", "Test not performed", "Insufficient quant.",
    "repeat sweat chloride testing", "Please reorder testing and re",
    "The received specimen cannot"
  )
)
chco <- chco %>%
  group_by(PatientID) %>%
  filter(Demographics.Available == "TRUE") %>%
  # Get only the earliest test
  slice_min(Age.at.Test.in.Weeks) %>%
  select(
    PatientID, Name, Genotypes1, Genotypes2, Sex, Age.at.Test.in.Years,
    Na, K, Cl, Na_Left, Na_Right, K_Left, K_Right, Cl_Left, Cl_Right
  ) %>%
  rename(Age = Age.at.Test.in.Years) %>%
  mutate(Name = tolower(gsub("[[:punct:] ]+", "", Name)))
# Fix CHCO diagnoses per Angela
chco_diagnoses <- read.csv("./Data_Cleaned/chco_diagnoses.csv")
chco <- full_join(chco, chco_diagnoses, by = join_by(PatientID))
chco$PatientID <- NULL
# For "LESS THAN 10", replace with half the limit of detection
chco[chco == "LESS THAN 10"] <- "5"
chco[, c("Cl", "Cl_Left", "Cl_Right")] <-
  lapply(chco[, c("Cl", "Cl_Left", "Cl_Right")], as.numeric)
# Import NJH and add diagnoses from Katie
njh <- read.csv("./Data_Cleaned/njh_full_list.csv",
  na.strings = c(
    "", "Test not performed", "Insufficient quant.",
    "repeat sweat chloride testing"
  )
)
njh_diagnoses <- read.csv("./Data_Cleaned/njh_diagnoses.csv", na.strings = "")
njh_diagnoses <- njh_diagnoses %>% filter(!is.na(Diagnosis))
njh <- full_join(njh, njh_diagnoses)
# Filter out those with no demo data, rename columns, etc.
njh <- njh %>%
  group_by(Patient.ID) %>%
  filter(Demographics.Available == "TRUE") %>%
  # Get only the earliest test
  slice_min(Age.at.test..in.years.) %>%
  select(
    Patient.ID, Patient.Name, Genotype.1, Genotype.2, Sex, Age.at.test..in.years.,
    Sweat.sodium, Sweat.potassium, Sweat.chloride, Sweat.sodium.left.arm,
    Sweat.sodium.right.arm, Sweat.potassium.left.arm, Sweat.potassium.right.arm,
    Sweat.chloride.left.arm, Sweat.chloride.right.arm, Diagnosis
  ) %>%
  rename(
    Name = Patient.Name, Genotypes1 = Genotype.1,
    Genotypes2 = Genotype.2, Age = Age.at.test..in.years., Na = Sweat.sodium,
    K = Sweat.potassium, Cl = Sweat.chloride, Na_Left = Sweat.sodium.left.arm,
    Na_Right = Sweat.sodium.right.arm, K_Left = Sweat.potassium.left.arm,
    K_Right = Sweat.potassium.right.arm, Cl_Left = Sweat.chloride.left.arm,
    Cl_Right = Sweat.chloride.right.arm
  ) %>%
  mutate(Name = tolower(gsub("[[:punct:] ]+", "", Name)))
njh$Patient.ID <- NULL
# Row bind, fill diagnosis, then remove dupes
df <- rbind(chco, njh)
df <- df %>%
  group_by(Name) %>%
  fill(Diagnosis) %>%
  distinct() %>%
  select(Name, Diagnosis, everything()) %>%
  arrange(Name, Age)
# Use each person's first sweat test
df <- df %>%
  group_by(Name) %>%
  slice_min(Age)
# If there are still duplicates, exclude them
dups <- df$Name[duplicated(df$Name)]
df <- df[!df$Name %in% dups, ]
# For those with left and right, average them
df$Na_mean <- rowMeans(df[, c("Na_Left", "Na_Right")])
df$Na <- coalesce(df$Na, df$Na_mean)
df$Cl_mean <- rowMeans(df[, c("Cl_Left", "Cl_Right")])
df$Cl <- coalesce(df$Cl, df$Cl_mean)
df$K_mean <- rowMeans(df[, c("K_Left", "K_Right")])
df$K <- coalesce(df$K, df$K_mean)
# Get rid of unnecessary columns, remove those with no diagnosis
df <- df %>%
  select(Name, Diagnosis, Genotypes1, Genotypes2, Sex, Age, Na, K, Cl) %>%
  filter(!is.na(Diagnosis))
# Set table options
mycontrols <-
  tableby.control(numeric.stats = c("Nmiss", "meansd", "medianq1q3", "range"))
# Save this for Angela et al. to check
write.csv(df,
  file = "./Data_Cleaned/sweat_variability_analysis_dataset.csv",
  row.names = F, na = ""
)
```

@@

- Imported re-identified CHCO data from Elin (5,046 rows and 4,748 unique IDs)
- Removed rows with `Demographics Available` not equal to `TRUE` and selected each person's earliest test by age.
  - This resulted in 3,123 rows with 3,084 unique IDs.
- Merged in the diagnoses from Angela.
- Values below the limit of detection were replaced with half the limit (e.g. "LESS THAN 10" was replaced with 5). Approximately 5% of the values were below the L.O.D.

>>>>

```{r setup}
#| include: false
library(Hmisc)
library(lubridate)
library(readxl)
library(tidyverse)
library(arsenal)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Variability",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Edith Zemanick/Sweat Variability"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| message: false
# Import CHCO, filter out those with no demo data, rename columns, etc.
chco <- read.csv("./Data_Cleaned/chco_reidentified.csv",
  na.strings = c(
    "", "Test not performed", "Insufficient quant.",
    "repeat sweat chloride testing", "Please reorder testing and re",
    "The received specimen cannot"
  )
)
chco <- chco %>%
  group_by(PatientID) %>%
  filter(Demographics.Available == "TRUE") %>%
  # Get only the earliest test
  slice_min(Age.at.Test.in.Weeks) %>%
  select(
    PatientID, Name, Genotypes1, Genotypes2, Sex, Age.at.Test.in.Years,
    Na, K, Cl, Na_Left, Na_Right, K_Left, K_Right, Cl_Left, Cl_Right
  ) %>%
  rename(Age = Age.at.Test.in.Years) %>%
  mutate(Name = tolower(gsub("[[:punct:] ]+", "", Name)))
# Fix CHCO diagnoses per Angela
chco_diagnoses <- read.csv("./Data_Cleaned/chco_diagnoses.csv")
chco <- full_join(chco, chco_diagnoses, by = join_by(PatientID))
chco$PatientID <- NULL
# For "LESS THAN 10", replace with half the limit of detection
chco[chco == "LESS THAN 10"] <- "9"
chco[, c("Cl", "Cl_Left", "Cl_Right")] <-
  lapply(chco[, c("Cl", "Cl_Left", "Cl_Right")], as.numeric)
# Import NJH and add diagnoses from Katie
njh <- read.csv("./Data_Cleaned/njh_full_list.csv",
  na.strings = c(
    "", "Test not performed", "Insufficient quant.",
    "repeat sweat chloride testing"
  )
)
njh_diagnoses <- read.csv("./Data_Cleaned/njh_diagnoses.csv", na.strings = "")
njh_diagnoses <- njh_diagnoses %>% filter(!is.na(Diagnosis))
njh <- full_join(njh, njh_diagnoses)
# Filter out those with no demo data, rename columns, etc.
njh <- njh %>%
  group_by(Patient.ID) %>%
  filter(Demographics.Available == "TRUE") %>%
  # Get only the earliest test
  slice_min(Age.at.test..in.years.) %>%
  select(
    Patient.ID, Patient.Name, Genotype.1, Genotype.2, Sex, Age.at.test..in.years.,
    Sweat.sodium, Sweat.potassium, Sweat.chloride, Sweat.sodium.left.arm,
    Sweat.sodium.right.arm, Sweat.potassium.left.arm, Sweat.potassium.right.arm,
    Sweat.chloride.left.arm, Sweat.chloride.right.arm, Diagnosis
  ) %>%
  rename(
    Name = Patient.Name, Genotypes1 = Genotype.1,
    Genotypes2 = Genotype.2, Age = Age.at.test..in.years., Na = Sweat.sodium,
    K = Sweat.potassium, Cl = Sweat.chloride, Na_Left = Sweat.sodium.left.arm,
    Na_Right = Sweat.sodium.right.arm, K_Left = Sweat.potassium.left.arm,
    K_Right = Sweat.potassium.right.arm, Cl_Left = Sweat.chloride.left.arm,
    Cl_Right = Sweat.chloride.right.arm
  ) %>%
  mutate(Name = tolower(gsub("[[:punct:] ]+", "", Name)))
njh$Patient.ID <- NULL
# Row bind, fill diagnosis, then remove dupes
df <- rbind(chco, njh)
df <- df %>%
  group_by(Name) %>%
  fill(Diagnosis) %>%
  distinct() %>%
  select(Name, Diagnosis, everything()) %>%
  arrange(Name, Age)
# Use each person's first sweat test
df <- df %>%
  group_by(Name) %>%
  slice_min(Age)
# If there are still duplicates, exclude them
dups <- df$Name[duplicated(df$Name)]
df <- df[!df$Name %in% dups, ]
# For those with left and right, average them
df$Na_mean <- rowMeans(df[, c("Na_Left", "Na_Right")])
df$Na <- coalesce(df$Na, df$Na_mean)
df$Cl_mean <- rowMeans(df[, c("Cl_Left", "Cl_Right")])
df$Cl <- coalesce(df$Cl, df$Cl_mean)
df$K_mean <- rowMeans(df[, c("K_Left", "K_Right")])
df$K <- coalesce(df$K, df$K_mean)
# Get rid of unnecessary columns, remove those with no diagnosis
df <- df %>%
  select(Name, Diagnosis, Genotypes1, Genotypes2, Sex, Age, Na, K, Cl) %>%
  filter(!is.na(Diagnosis))
# Set table options
mycontrols <-
  tableby.control(numeric.stats = c("Nmiss", "meansd", "medianq1q3", "range"))
# Save this for Angela et al. to check
write.csv(df,
  file = "./Data_Cleaned/sweat_variability_analysis_dataset.csv",
  row.names = F, na = ""
)
```

@@

- Imported re-identified CHCO data from Elin (5,046 rows and 4,748 unique IDs)
- Removed rows with `Demographics Available` not equal to `TRUE` and selected each person's earliest test by age.
  - This resulted in 3,123 rows with 3,084 unique IDs.
- Merged in the diagnoses from Angela.
- Values below the limit of detection were replaced with half the limit (e.g. "LESS THAN 10" was replaced with 9). Approximately 5% of the values were below the L.O.D.
